Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-07-30 Scripts/Manuscript_Draft.Rmd
Sent revisions to co authors

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
subtitle: "CORECTIONS ARE IN BOLD. DO NOT UNBOLD"
# subtitle: "Most neighbor nations share marine exploited species"
output:
  pdf_document:
    fig_caption: yes
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

@@

# Reviewing issues

## General comments
- Not sure I understand/respond properly to Rev 3. Lines 58-59.

## Questions for Gabs:
- Teh et al comment. Do we include another data or just the explanation is fine?
- **Rev 1.** Line 132: Why? These last sentence is hard to understand.

## Figures
- Fix figure 3 solomon islands

## Analysis
- Convert revenue to 2019 real USD

## To editor
- Note that revenue has increased substantially. this is because we converted the 2010 real USD to 2018, as suggested by Rev. #3

## Specific comments
- **Rev 1.** Line 305 to 320:

@@

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 65 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear **more intensively fished** than those mainly catching non-transboundary species. Our study has direct implications **for fisheries management** of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

@@

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 241 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level **(or even as a group of species) like the case of shrimps (e.g. Litopenaeus stylirostris or Farfantepenaeus californiensis) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt], hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf].** Also, recent research has suggested strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis **while still making it relevant from a management perspective**. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only considered a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting transboundary species within EEZs **caught** an **annual** average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating **a yearly average of USD 65,483 million in fishing revenue during the same time period (Fig. 1)**. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 30,700 million in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. **This result suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated. However, it is important to mention that the referenced article reports yearly FAO reported catch data for both EEZs and the high seas up to 2006. Thus, the divergence between studies can be influenced by the data source and time period in addition to the higher amount of transboundary species identified in the present study. ... Using the FAO reported data for 2006, the average participation of the identified transboundary species was 61%, still higher than that reported by Teh and Sumaila (2015).**

Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, we found that Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile [@Cashion:2018cg], and thus discrete for Ecuador (**spatial threshold < 10%**). The **annual average contribution of discrete species (in our previous example would be only considering Ecuador's anchoveta catch) between 2005 and 2014** was 14.8 million tons representing USD 22,167 million in fishing revenue **over the same time period**. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary **according to our methods of determination (See methods)**. These species represented only **an annual average of** 2.8 million tons and 3,691 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

@@

**We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is not to infer stock status but to show the catch trend of transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods).** We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, transboundary species that are only fished by one country are generally less likely to be over-exploited compared to those shared by neighboring countries (Table S2). Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052). This result suggests that the management of discrete species could be better off as the variation between Categories A and B is closer (*p* = 0.052) than for transboundary species (*p* = 0.001).

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that contributed an average of 598.4 thousand tons to **total** catches (8.19% of total) and USD 850.2 million in **fishing** revenue (7.17% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change in the distribution of share species is key to develop local adaptation methods to anticipate negative impacts to sustainability. **For example, understanding how climate change will modify the proportion that each neighboring EEZ has over a sharing species [@Palacios-Abrantes_OC], when this change is going to happen, and the economic consequence of such effects can make international treaties more resilient to climate change [@Sumaila_oc].**

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Total catch between 2005 and 2014 of top five fishing countries capturing the top five shared species of each ecosystem preference (shown in bold on A). Note that there could be > 5 species due to similar sharing values. **Other consist in species that have no classification according to FishBase.**](../Figures/Fig4.png)

**However, this category also includes species that are not considered highly migratory but still shared by numerous neighboring EEZs (e.g. Belone belone). Species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm] but the potential boundary-crossing of these species must not be ignored. In addition, the world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. Despite coral reef-related species being among the less shared neighboring EEZs identified in this study (Fig. 4), coral regions, often share multiple species through larvae connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. In some areas like the Caribbean and the Western Indian Ocean, transboundary marine protected areas have been identified as potential tools for achieving fisheries and conservation goals [@Levin:2018in; Perez:2019dm].** The proper management of coral reef species is very important for many coastal communities, particularly in developing regions, that are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. **4B**) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc].

@@

Our study identified species currently shared by neighboring coastal nations and highlights the importance of these species' contribution to global capture and revenue from wild fisheries. Our results show that captures and revenues from shared species are substantially higher than those from discrete species, as well as than previously estimated **[@Teh:2015gd]**. This result highlights the importance of transboundary fisheries and their potential contribution to food security and wellbeing. Moreover, we show significant differences in **the catch trends** of transboundary and discrete species, suggesting a need for improving the management of transboundary fisheries. Previous work has found that collaboration will result in better outcomes for shared marine resources [@Munro:2002uf]. Thus, countries should pay specific attention in the management of the identified transboundary stocks potentially creating collaborative management plans. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability, and can be flexible to accommodate the uncertain future driven by climate change.

@@

The previous models combine observational data with a series of biotic and abiotic data to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated [@Zeller2016]. **Roughly, the Sea Around Us method takes each country's officially reported catch data (e.g. National, FAO or RFMO). Secondly, it uses literature (e.g. peer review, grey literature) to identify missing components (e.g. species, gears) and and sources of alternative information for missing components. It then derives country estimates for missing data and creates time series interpolation. Finally the estimated and official data are aggregated for a total reconstructed catch data See [@Pauly:2019bf] and [@Zeller2016] for catch reconstruction and spatial allocation details.** We used the *Sea Around Us* catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. We selected this time frame to investigate the recent (our time frame includes the **last decade of available data**) contribution of transboundary fisheries to catches and revenue from fisheries and to reduce the uncertainty embedded in the reconstruction process (See key uncertainties, below).

@@

As previously mentioned, we define transboundary species as those marine species that occur within the EEZs of two or more neighboring countries. Hence, according to this criteria a species was only considered as transboundary if it was shared between two neighboring countries, regardless of the species extended distribution. We defined EEZ boundaries using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides EEZs by regions (e.g., Mexico Pacific and Mexico Atlantic), and determined the intersections between polygons using the R package sf. When estimating transboundary species, we filtered out those shared by EEZs sub-regions, and when aggregating results by country, species that occurred in more than one sub-region were only accounted for once. Species that were present in EEZs that were non-continental territories neighboring other countries were kept separately (e.g., Argentina and Falkland Islands), but removed in cases where the non-continental territory belonged to the same nation (e.g., Brazil and Fernando de Noronha). Associated states like Puerto Rico and Hong Kong were added to their Sovereign (e.g. Puerto Rico with the United States and Hong Kong with China).

>>>>

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
subtitle: "CORECTIONS ARE IN BOLD. DO NOT UNBOLD"
# subtitle: "Most neighbor nations share marine exploited species"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

@@

@@

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 75 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear **more intensively fished** than those mainly catching non-transboundary species. Our study has direct implications **for fisheries management** of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

@@

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 243 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level **(or even as a group of species) like the case of shrimps (e.g. Litopenaeus stylirostris or Farfantepenaeus californiensis) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt], hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf].** Also, recent research has suggested strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis **while still making it relevant from a management perspective**. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only consider a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the 938 species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting transboundary species within EEZs **caught** an **annual** average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating **a yearly average of USD 75,961 million in fishing revenue during the same time period**. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 35,612 million (**converted to in 2019 USD**) in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. **Such difference could be due to the utilization of reported catch data by Teh and Sumaila instead of reconstructed data and a time period difference (Teh and Sumaila's study ends in 2006). To compare the two studies, we used the FAO global reported data identified at the species level by the *Sea Around Us* and found that in 2006, the 633 transboundary species accounted for 40.4 million tons. While this is not necessarily a direct comparison (e.g. methodology differences might still exist), it suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated due to an incomplete knowledge on the transboundary nature of marine fished species**

Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, our study shows that Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile and thus discrete for Ecuador (**spatial threshold between Ecuador and Peru < 10%**) **[@Cashion:2018cg]**. The **annual average contribution of discrete species (in our previous example would be only considering Ecuador's anchoveta catch) between 2005 and 2014** was 10.8 million tons representing USD 19,243 million in fishing revenue **over the same time period**. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary **according to our methods of determination (See methods)**. These species represented only **an annual average of** 2.8 million tons and 4,282 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

@@

**We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is not to infer stock status but to show the catch trend of transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods).** We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, **occasions where a transboundary species is only fished by one country are generally less likely to have a decreasing catch trend compared to when the species is shared by neighboring countries (Table S2)**. Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052).

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that **straddles into the high seas** that contributed an average of 598.4 thousand tons to **total** catches (8.19% of total) and USD 850.2 million in **fishing** revenue (7.17% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change in the distribution of shared species is key to develop local adaptation methods to anticipate negative impacts to sustainability. **For example, understanding how climate change will modify the proportion that each neighboring EEZ has over a sharing species [@PalaciosAbrantes:2020vt], when this change is going to happen, and the economic consequence of such effects can make international treaties more resilient to climate change [@RashidSumaila:tp].**

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Total catch between 2005 and 2014 of top five fishing countries capturing the top five shared species of each ecosystem preference (shown in bold on A). Note that there could be > 5 species due to similar sharing values. **Other consist of species that have no classification according to FishBase.**](../Figures/Fig4.png)

**However, this category also includes species that are not considered highly migratory but still shared by numerous neighboring EEZs (e.g. Belone belone). Species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm] but the potential boundary-crossing of these species must not be ignored. In addition, the world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. Despite coral reef-related species being among the less shared neighboring EEZs identified in this study (Fig. 4), coral regions often share multiple species through larvae connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. In some areas like the Caribbean and the Western Indian Ocean, transboundary marine protected areas have been identified as potential tools for achieving fisheries and conservation goals [@Levin:2018in; @Perez:2019dm].** The proper management of coral reef species is very important for many coastal communities, particularly in developing regions, that are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. **4B**) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc].

@@

Our study identified species currently shared by neighboring coastal nations and highlights the importance of these species' contribution to global capture and revenue from wild fisheries. Our results show that captures and revenues from shared species are substantially higher than those from discrete species, as well as than previously estimated**[@Teh:2015gd]**. This result highlights the importance of transboundary fisheries and their potential contribution to food security and wellbeing. Moreover, we show significant differences in **the catch trends** of transboundary and discrete species, suggesting a need for improving the management of transboundary fisheries. Previous work has found that collaboration will result in better outcomes for shared marine resources [@Munro:2002uf]. Thus, countries should pay specific attention in the management of the identified transboundary stocks potentially creating collaborative management plans. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability, and can be flexible to accommodate the uncertain future driven by climate change.

@@

The previous models combine observational data with a series of biotic and abiotic data to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated [@Zeller2016]. **Roughly, the Sea Around Us method takes each country's officially reported catch data (e.g. National, FAO or RFMO). Secondly, it uses literature (e.g. peer review, grey literature) to identify missing components (e.g. species, gears) and sources of alternative information for missing components. It then derives country estimates for missing data and creates time series interpolation. Finally the estimated and official data are aggregated for a total reconstructed catch data See [@Pauly:2019bf] and [@Zeller2016] for catch reconstruction and spatial allocation details.** We used the *Sea Around Us* catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. We selected this time frame to investigate the recent (our time frame includes the **last decade of available data**) contribution of transboundary fisheries to catches and revenue from fisheries and to reduce the uncertainty embedded in the reconstruction process (See key uncertainties, below).

@@

As previously mentioned, we define transboundary species as those marine species that occur within the EEZs of two or more neighboring countries. Hence, according to this criteria a species was only considered as transboundary if it was shared between two neighboring countries, regardless of the species extended distribution. We defined EEZ boundaries using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org), **noting that it subdivides EEZs by 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories**, and determined the intersections between polygons using the R package *sf* [@PackagesfSimpleF:2018vp]. When estimating transboundary species, we filtered out those shared by EEZs sub-regions (e.g. USA Gulf of Mexico and USA Atlantic), and when aggregating results by country, species that occurred in more than one sub-region were only accounted for once. Species that were present in EEZs that were non-continental territories neighboring other countries were kept separately (e.g., Argentina and Falkland Islands), but removed in cases where the non-continental territory belonged to the same nation (e.g., Brazil and Fernando de Noronha). Associated states like Puerto Rico and Hong Kong were added to their Sovereign (e.g. Puerto Rico with the United States and Hong Kong with China).
